---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# SICNN

<!-- badges: start -->

[![](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://lifecycle.r-lib.org/articles/stages.html#experimental) [![](https://img.shields.io/github/last-commit/aliaksah/SICNN.svg)](https://github.com/aliaksah/SICNN/commits/main) [![](https://img.shields.io/github/languages/code-size/aliaksah/SICNN.svg)](https://github.com/aliaksah/SICNN) [![R-CMD-check](https://github.com/aliaksah/SICNN/workflows/R-CMD-check/badge.svg)](https://github.com/aliaksah/SICNN/actions) [![License: MIT](https://img.shields.io/badge/license-MIT-blue.svg)](LICENSE)

<!-- badges: end -->

The goal of SICNN is to implement Smooth Information Criterion Neural Networks in R, using the torch package. SICNN is a scalable architecture designed exclusively for seamless, continuous parameter space pruning. Evolved from LBBNNs, SICNN operates dynamically and directly on model weight structures by natively leveraging a continuous smooth approximation of the $L_0$ norm.

## Installation

You can install the development version of SICNN from [GitHub](https://github.com/aliaksah/SICNN) with:

``` r
# install.packages("pak")
pak::pak("aliaksah/SICNN")
```

## Example

This example demonstrates how to implement a SICNN model on the raisin dataset.

```{r example}
library(SICNN)
library(ggplot2)
library(torch)

# the get_dataloaders function handles preprocessing
loaders <- get_dataloaders(Raisin_Dataset, train_proportion = 0.8, train_batch_size = 720, test_batch_size = 180)
train_loader <- loaders$train_loader
test_loader <- loaders$test_loader

```

To initialize the SICNN, we need to define the type of problem and network dimension sizes.

```{r}
problem <- 'binary classification'
sizes <- c(7,5,5,1) # 7 input variables, two hidden layers of 5 neurons each, 1 output neuron.
device <- 'cpu' # can also be mps or gpu.
```

We define the core structural model using `SICNN_Net`.

```{r}
torch_manual_seed(0)
model_input_skip <- SICNN_Net(problem_type = problem, sizes = sizes, 
                              input_skip = TRUE, device = device)
```

To train the models, we apply `train_SICNN` which relies on an exponentially decaying smooth L0 surrogate to effectively drop inactive pathways.

```{r,results='hide'}
# Track the number of samples mapping to log(n) regularization
n_train <- nrow(Raisin_Dataset) * 0.8
results_input_skip <- suppressMessages(train_SICNN(
    epochs = 5000, restarts = 1, SICNN = model_input_skip, 
    lr = 0.01, train_dl = train_loader, device = device,
    scheduler = "step", sch_step_size = 2000, n_train = n_train,
    epsilon_1 = 1, epsilon_T = 1e-4, steps_T = 4000, sic_threshold = 0.5,
    penalty = 15 * log(n_train)
))
```

Evaluate structural performance and accuracy on validation samples natively:

```{r}
validate_SICNN(SICNN = model_input_skip, num_samples = 1, test_dl = test_loader, device=device)
```

Plot the global interconnected sparse structure of the active model pathways:

```{r}
plot(model_input_skip, type = 'global', vertex_size = 13, edge_width = 0.6, label_size = 0.6)
```

This structural architecture reflects precisely which of the 7 initial parameters remain viable variables. We can also plot local interactions relative to one patient/run directly via input-skip activations:

```{r}
x <- train_loader$dataset$tensors[[1]] 
data <- x[42,] # Evaluate specific dataset slice dynamically
plot(model_input_skip, type = 'local', data = data)
```
