---
title: "Getting started with SICNN"
output:
  rmarkdown::html_vignette:
    df_print: paged
params:
  eval: false
vignette: >
  %\VignetteIndexEntry{Getting started with SICNN}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = params$eval
)
```

## Introduction

SICNN implements Smooth Information Criterion Neural Networks in R using the torch framework. This vignette walks through basic usage: data preparation, model definition, training, validation, and visualization.

## Setup

```{r}
library(SICNN)
library(ggplot2)
library(torch)
```

## Data loaders

```{r}
loaders <- get_dataloaders(Raisin_Dataset, train_proportion = 0.8,
                           train_batch_size = 720, test_batch_size = 180)
train_loader <- loaders$train_loader
test_loader  <- loaders$test_loader
```

## Define the model

```{r}
problem <- 'binary classification'
sizes <- c(7,5,5,1)
device <- 'cpu'

torch_manual_seed(0)

# We use the SICNN_Net structure natively
model_input_skip <- SICNN_Net(problem_type = problem, sizes = sizes, 
                              input_skip = TRUE, device = device)
```

## Train

```{r}
n_train <- nrow(Raisin_Dataset) * 0.8
# Train the model mapping the objective to the Smooth BIC penalty naturally
results_input_skip <- train_SICNN(epochs = 5000, restarts=1, SICNN = model_input_skip,
                                  lr = 0.01, train_dl = train_loader, device = device,
                                  scheduler = "step", sch_step_size = 500, n_train=n_train,
                                  epsilon_1 = 1, epsilon_T = 1e-5, steps_T = 500, 
                                  sic_threshold = 0.5, penalty = 5 * log(n_train))
```

## Validate

```{r}
validate_SICNN(SICNN = model_input_skip,
               test_dl = test_loader, device = device)
```

## Plot structure and explanations

```{r}
# Plot sparse global architecture
plot(model_input_skip, type = "global", vertex_size = 10, edge_width = 0.4, label_size=0.3)

# Plot local explanation on an individual
x <- torch::dataloader_next(torch::dataloader_make_iter(train_loader))[[1]]
inds <- sample.int(dim(x)[1], 1)
data <- x[inds,]
plot(model_input_skip, type = "local", data = data)

# Plot local explanation with Delta method uncertainty (using block-diagonal Fisher)
plot(model_input_skip, type = "local", data = data, 
     uncertainty = TRUE, fisher_dataloader = test_loader,
     covariance_type = "block-diagonal")
```

Note: All chunks in this vignette are non-evaluated by default to ensure fast builds and avoid backend constraints on CRAN/CI. You can set `eval=TRUE` locally.
