---
title: "Tutorial 4: Convolutional Architecture on KMNIST"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Tutorial 4: Convolutional Architecture on KMNIST}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = FALSE
)
```

## Introduction
This tutorial demonstrates how to leverage the `SICNN` framework for a Convolutional Neural Network on the Kuzushiji-MNIST (KMNIST) dataset. Since SICNN modules (`SICNN_Conv2d` and `SICNN_Linear`) track `weight_mean` and `alpha_active_path` identically, they can be composed into arbitrary deep learning architectures like the `SICNN_ConvNet`.

## Loading Dependencies and the Dataset

```{r setup, eval=FALSE}
# install.packages("torchvision") 
library(torch)
library(torchvision)
library(SICNN)

dir <- "./dataset/kmnist"   #directory to install dataset
train_ds <- torchvision::kmnist_dataset(
  dir,
  download = TRUE,
  transform = torchvision::transform_to_tensor)

test_ds <- torchvision::kmnist_dataset(
  dir,
  train = FALSE,
  transform = torchvision::transform_to_tensor)

train_loader <- torch::dataloader(train_ds, batch_size = 100, shuffle = TRUE)
test_loader <- torch::dataloader(test_ds, batch_size = 100)
```

## Initializing the Convolutional Architecture

We use the native `SICNN_Conv2d` layers to construct the feature extractor, and standard `SICNN_Linear` layers for the classifier. The built-in `SICNN_ConvNet` module wraps these layers specifically for MNIST-shaped classification tasks.

```{r conv-init, eval=FALSE}
device <- 'cpu' # Can be swapped with 'gpu' or 'mps' on supported hardware
torch::torch_manual_seed(42)

conv_layer_1 <- SICNN_Conv2d(in_channels = 1, out_channels = 32, kernel_size = 5, device = device)
conv_layer_2 <- SICNN_Conv2d(in_channels = 32, out_channels = 64, kernel_size = 5, device = device)

# After two max pools of 2x2, the 28x28 image becomes 4x4.
# (28 - 4)/1 = 24. 24/2 = 12.
# (12 - 4)/1 = 8. 8/2 = 4. 4*4*64 = 1024
linear_layer_1 <- SICNN_Linear(in_features = 1024, out_features = 300, device = device, bias = FALSE)
linear_layer_2 <- SICNN_Linear(in_features = 300, out_features = 10, device = device, bias = FALSE)


model <- SICNN_ConvNet(conv1 = conv_layer_1, 
                       conv2 = conv_layer_2, 
                       fc1 = linear_layer_1, 
                       fc2 = linear_layer_2, 
                       device = device)
                       
model$to(device = device)
```

## Training the Model

The framework exposes the `train_SICNN` functionality seamlessly to the novel architecture since `SICNN_ConvNet` inherits and defines standard SIC metrics (like `smooth_param_count`).

```{r train, eval=FALSE}
# We use epochs = 3 for this demonstration
train_SICNN(epochs = 3, 
            SICNN = model, 
            lr = 0.001, 
            train_dl = train_loader, 
            n_train = length(train_ds),
            device = device)
```

## Validation and Predictions

Once the model has converged, we calculate standard test-set predictions via multi-draw validation to understand its epistemic uncertainty and overall sparsity.

```{r validate, eval=FALSE}
validate_SICNN(model, num_samples = 10, test_dl = test_loader, device = device)

# Predict draws on the test set explicitly
predictions <- predict(model, newdata = test_loader, draws = 10)
cat(sprintf("Prediction tensor dimensions: %s\n", paste(dim(predictions), collapse="x")))
print(predictions[, 1, ])
```
